Discrete Quantum Theories and Computing

### 2019-01-21 dissertation.tex
dissertation.lyx Exported to HTML and grammar checked by Grammarly.

\begin{acknowledgments*}\todo{Confirm people to be acknowledged
like the way being acknowledged\ldots{} }I want to thank Prof.~Gerardo
Ortiz, Prof.~Amr Sabry (\textarabic{عمرو صبري}), and Prof.~Andrew
Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2018}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on ``Geometry of discrete quantum computing''~\cite{geometry2013}
and ``Discrete quantum theories'' (DQT)~\cite{DQT2014}; most of
Chapter~\ref{chap:QIVPM} is based on ``Quantum Interval-Valued
Probability: Contextuality and the Born Rule'' (QIVPM)~\cite{THOS2018}.
Together with Prof.~Dylan Thurston and Prof.~Shouhong Wang (汪守宏),
I want to thank you for getting together to understand what I did
and passing my dissertation proposal Spring 2017. I also thank Prof.\ Dylan
Thurston for serving in my advisory committee and my Tier 3 committee,
with other advisory committee members passing my Computer Science
qualifying exam Spring 2016. Especially, my survey of real computation
in the qualifying exam inspired me the comparison of quantum computer
and analog computer in Chapter~\ref{chap:Introduction}. Chapter~\ref{chap:Introduction}
also contains the motivation in my application for Rethinking Foundations
of Physics 2017 workshop reviewed by Prof.\ Amr Sabry~(\textarabic{عمرو صبري}).
I want to thank Prof.\ Lawrence Moss for inviting me to present our
results in the interdisciplinary logic seminar and theory seminar
in Computer Science, co-chairing my Tier 3 committee, writing an assessment
letter for me, and drawing my attention on Prof.~Abramsky's paper~\cite{Abramsky2012},
which inspired me to merge quantum probability with discrete quantum
theories. And I thank John Gardiner for inspiring discussion on the
difficulties merging them~\cite{Gardiner2014}. These results are
improved and motivated the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{subsec:Discrete-Deutsch-algorithm}.
I want to thank Traci Nagle for going through the literature review
about finite precision measurement and contextuality discussed for
CSCI-Y790 Writing and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.
I also thank Elizabeth ``Betsy'' Merceron and Kexin ``Casey''
Chen (谌可心) for teaching me English in general so that others can understand
me more easily.

\todo{Type a table for acknowledgment so that I would have to write
``I thank'' so many times here? }Many people in both department
of Mathematics and Computer Science need to do something more to make
my double-major works smoothly. I need to thank all of them, especially
directors of graduate study in Mathematics, especially Prof.\ Christopher
Judge and Prof.\ Matthias Weber; directors of PhD. study in Computer
Science, especially Prof.\ Yuqing Melanie Wu (吴愈青) and Prof.\ \textturkish{Funda
Ergun}.

@@

Thank all my friends for your supports, including:

@@

\item Hao-Chun Lee (李浩君) and Hsien-Ching Kao (高憲慶) for discussing
what I did in the prospective of a general PhD. in Physics and working
later in more computational industry.

@@

\begin{abstract*}Most quantum computing models are based on the continuum
of real numbers, while classical digital computers faithfully realize
only discrete computational models. Analog computers appear to be
an option, but in reality are far weaker than would be needed for
computational models requiring real numbers. One approach to resolving
this conflict is to find consistent mathematical ways to limit measurement
precision to computable contexts that do not require incomputable
real numbers. Our goal is to build a more philosophically consistent
models by investigating discrete quantum computing using finite number
systems, and, alternatively, by incorporating finite precision measurement
using intervals into quantum theory.

@@

Finally, we shift our attention to consider quantum interval-valued
probability measures (IVPMs), which potentially embody both finite
precision measurement and a sensible correspondence to standard quantum
probability. This interval-valued framework not only provides a natural
generalization of both classical IVPMs and conventional quantum probability
measures, but also allows us to establish bounds on the validity of
the Kochen-Specker and Gleason theorems in realistic experimental
environments.\end{abstract*}

@@

Since we as human cannot distinguish quantities differing only by
an arbitrarily small extent \cite{Turing_1937,Gisin2017}, we want
to incorporate this limitation into our quantum mechanical model to
better capture our ability of predicting quantum phenomenon and the
power of realistic quantum computers. Since we are agnostic about
whether the reality is ultimately discrete or continuous, we considered
two types of quantum models. By assuming the quantum states are ultimately
discrete and distinguishable, our first type of quantum theories replaces
the complex numbers by finite fields. We then focus on the limitation
of distinguishability itself, incorporate the finite precision measurement
into our model, and consider quantum interval-valued probability measures.

After first suggested by Richard Feynman 40 years ago~\cite{Feynman1982Simulating},
IBM, \textitalian{Rigetti}, Google, Microsoft, and many other companies
and countries start to build realistic quantum computers recently
\cite{Shieber2018,Kahn2019}, and seek more efficient ways to simulate
chemical molecules and to break down RSA \cite{boneh1999twenty,wiki:RSAFactoring}.
No matter how to, classical or quantum, attack these two tasks, they
require two different types of quantities. On one hand, a question
and its answer of RSA are both discrete integers. On the other hand,
molecules are described by quantities including the domain of their
position wave functions and the probability amplitudes of these wave
functions, which look like continuous at first sight. However, either
these quantities might later be discovered to be discrete, or their
exact values may never be pinned down. Hence, modeling quantum computing
based on the exact values of continuous quantities cannot really describe
the computational power of a realistic quantum computer. To better
understand this issue without struggling with the quantum theory,
it is instructive to review how classical quantities are characterized
and computed.

@@

In contrast, the non-apparently discrete quantities confused many
cultures since very beginning. For example, the ancient Chinese philosopher
Chuang-tzu (庄子) didn't believe these quantities have a minimum unit,
and expressed ``If a one-foot-long stick is cut into halves every
day, the cutting will never come to an end''\footnote{``一尺之棰，日取其半，萬世不竭。'' in Chinese \cite{Zhuangzi1999,Liu2018}.}.
As another example, Greek mathematicians in the school of Pythagoras
believed every mathematical model must be ultimately discrete, and
they in legend murdered Hippasus because he proved there is no minimum
unit between a side of a square and its diagonal. Comparing the idea
that the length of a square's diagonal is not a physical quantity
or there is no square physically, working with a mathematical model
without minimum unit seems more appealing. Since the mathematical
models are continuous, their corresponding physical quantities are
naturally assumed to be continuous. Later, these continuous quantities
become the basic component of analog computers. After logarithm was
discovered, a slide rule, also known as a slipstick, is used to compute
multiplication, division, and more complex operations as a mechanical
analog computer~\cite{wiki:SlideRule}. Charge amplifiers are used
to build electrical analog computers and compute the integration in
calculus \cite{wiki:Integrator,wiki:ChargeAmplifier}.

Although it is convenient to assume physical quantities is as continuous
as their mathematical models, each physical quantity discussed in
the previous paragraph could not have a simple one-to-one correspondence
with a real number. Chuang-tzu's stick are made by molecules and cannot
be cut in halves endlessly. The input charge of an integrator must
be an integer multiple of the elementary charge. A prefect square
cannot be drawn physically. Even we don't draw it, but merely ask
whether it exists in the space. This question is still beyond our
ability to answer because whether the space itself is ultimately discrete
or continuous is still an unsolved question. Even if a prefect square
exists physically, we may not able to decide whether the ``square''
we discussed is really prefect or almost prefect differing only by
an arbitrarily small extent. Similar situation applying to the length
of a slide rule, it is beyond our ability to decide whether its length
is ultimately discrete or continuous. Even if its length is continuous,
identifying which real number represents its length requires infinite
precision, but the precision in reality is limited generally three
or four significant figures~\cite{wiki:AnalogComputer}. As the results,
even if analog computers really store real numbers, they cannot be
precisely read, written, and used to branch the computation. The opposite
of the last statement is sometimes assumed by real computational models.
For example, Blum, Cucker, Shub, and Smale modeled their BCSS machine
to allow the operations of deciding whether a number is greater or
equal to zero over real numbers and whether a number is exactly zero
over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Hence, when the quantities used to branch the computation closes to
zero, the branch chosen by a BCSS machine cannot reliably predict
the branch chosen by a realistic analog computer. Since two branches
of computations may not have any relations, the follow-up computational
paths may be significantly different, and this difference is unlikely
to be compensated by error analysis techniques easily like the butterfly
effect. Therefore, the computational power of analog computers might
be far weaker than the prediction based on this kind of theoretical
models.%
\begin{comment}
\todo{Maybe extended to consider probabilistic model? }
\end{comment}

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How does the theorical model predict the behavior of the Physical device? & 
Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Like its classical counterparts, quantum circuit model, the
most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states. To address the similar inconsistency, we first review the
``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}, where the probability amplitudes are in the field of complex numbers.
If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, considering the precision
of quantities could still result in a better model. However, comparing
to the precision of probability amplitudes, it is easier to consider
the precision of their inducing probabilities because the idea of
``imprecise probabilities'' have well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }This idea will be extended
to quantum domain as quantum interval-valued probability measures
(QIVPM).

@@

Given a pure state~$\ket{\Phi}$, when we measure an observable represented
by an Hermitian matrix~$\mathbf{O}$, the measurement result is one
of the eigenvalues of $\mathbf{O}$. The probability of getting a
particular eigenvalue~$\lambda$ is $\melem{\Phi}{P}{\Phi}$, where
$P$ is the projection operator onto the eigenspace of $\lambda$.
This rule of computing the probability is called the Born rule \cite{Born1983,Mermin2007,Jaeger2007},
which is used when we want to extract information from a quantum computer.
For any mixed state~$\rho$, the generalized Born rule induces a
conventional quantum probability measure $\muB_{\rho}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projection operators on a given
Hilbert space. Conversely, any quantum probability measure $\mu:\events\rightarrow[0,1]$
can be induced from a mixed state $\rho$ in the Hilbert space of
dimension $D\ge3$ according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
In another word, this state $\rho$ is the unique state consistent
with any given quantum probability measure.
\end{comment}
The geometrical structure of states in CQT is first reviewed in Sec.~\ref{CQC1qubitBloch.sec},
where a pure state is represented as a vector in a Hilbert space,
and an irreducible state is actually a point in a projective Hilbert
space \cite{MosseriDandoloff2001,Jaeger2007} also known as a complex
projective space \cite{Hatcher2001,Bengtsson2007}. In Sec.~\ref{sec:fuzzy},
we will construct a quantum probability space based on classical probability
spaces, apply Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}
to recover the Born rule \cite{Born1983,Mermin2007,Jaeger2007}, and
define the expectation value of an observable. The expectation value
will then be used in Sec.~\ref{sec:The-geometry-of} to define purity
which provides an easy entanglement test.

\begin{comment}
In Section~\ref{modalquantum}, we examine previously-introduced
quantum theories defined over unrestricted finite fields and show
in Section~\ref{modalquantumcomputing} that this approach leads
to theories with such bizarre powers that they are probably nonphysical.
Although a version of quantum theory defined over a two-valued field
can express simple algorithms such as quantum teleportation, it is
so weak that it cannot express Deutsch's algorithm. This quantum theory
is, however, also so powerful that it can be used to solve an unstructured
database search of size $N$ using $O(\log(N))$ steps, which outperforms
the known asymptotic bound $O(\sqrt{N})$ in conventional quantum
computing.
\end{comment}
\begin{comment}
Our first discrete model replaces the complex numbers by an unrestricted
finite field~$\mathbb{F}_{q}$ \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
In this model, an $n$-qubit state is a non-zero vector in $\mathbb{F}_{q}^{2^{n}}$.
When we measure a given state~$\ket{\Phi}$, an observable is replaced
by a basis~$\mathcal{B}$ of $\mathbb{F}_{q}^{2^{n}}$. Since $\ket{\Phi}$
can be represented by the summation across a unique subset~$\mathcal{S}$
of the basis~$\mathcal{B}$, whether it is possible or impossible
to measure a basis vector~$\ket{i}$ depends on whether $\ket{i}$
is in $\mathcal{S}$ or not. Because this model only predicts whether
a result is possible or impossible, it is called the modal quantum
theory. Its computational model, called the modal quantum computing,
is far from conventional quantum computing. Although modal quantum
computing can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007}.
This quantum theory is, however, also so powerful that it can be used
to solve an unstructured database search of size $N$ using $O(\log(N))$
steps, which outperforms the known asymptotic bound $O(\sqrt{N})$
in conventional quantum computing \cite{Grover:1996:FQM:237814.237866,BennettBernsteinBrassardVazirani1997,Mermin2007,Jaeger2007}.
\end{comment}
In Chapter~\ref{chap:Unrestricted-Finite-Fields}, we examine previously-introduced
quantum theories defined over unrestricted finite fields \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
It is called the modal quantum theory because it can only predict
whether a measure result is possible or impossible, but not its probability.
Although a version of quantum theory defined over the two-valued field
can express simple algorithms such as quantum teleportation \cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007,Schumacher2012-SCHMQT},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007,DQT2014}.
This quantum theory is, however, also so powerful that it can be used
to efficiently solve the $\usat$ problem~\cite{Papadimitriou1993},
which is as hard as a general $\NP$-complete problem~\cite{Valiant198685}.

>>>>

\begin{acknowledgments*}\todo{Confirm people to be acknowledged
like the way being acknowledged\ldots{} }I want to thank Prof.~Gerardo
Ortiz, Prof.~Amr Sabry (\textarabic{عمرو صبري}), and Prof.~Andrew
Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2018}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on ``Geometry of discrete quantum computing''~\cite{geometry2013}
and ``Discrete quantum theories'' (DQT)~\cite{DQT2014}; most of
Chapter~\ref{chap:QIVPM} is based on ``Quantum Interval-Valued
Probability: Contextuality and the Born Rule'' (QIVPM)~\cite{THOS2018}.
Together with Prof.~Dylan Thurston and Prof.~Shouhong Wang (汪守宏),
I want to thank you for getting together to understand what I did
and passing my dissertation proposal Spring 2017. I also thank Prof.\ Dylan
Thurston for serving in my advisory committee and my Tier 3 committee,
with other advisory committee members passing my Computer Science
qualifying exam in Spring 2016. Especially, my survey of real computation
in the qualifying exam inspired me the comparison of quantum computer
and analog computer in Chapter~\ref{chap:Introduction}. Chapter~\ref{chap:Introduction}
also contains the motivation in my application for Rethinking Foundations
of Physics 2017 workshop reviewed by Prof.\ Amr Sabry~(\textarabic{عمرو صبري}).
I want to thank Prof.\ Lawrence Moss for inviting me to present our
results in the interdisciplinary logic seminar and theory seminar
in Computer Science, co-chairing my Tier 3 committee, writing an assessment
letter for me, and drawing my attention on Prof.~Abramsky's paper~\cite{Abramsky2012},
which inspired me to merge quantum probability with discrete quantum
theories. And I thank John Gardiner for inspiring discussion on the
difficulties merging them~\cite{Gardiner2014}. These results are
improved and motivated the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{subsec:Discrete-Deutsch-algorithm}.
I also thank Traci Nagle, Elizabeth ``Betsy'' Merceron, and Kexin
``Casey'' Chen (谌可心) for teaching me English in general so that
others can understand me more easily.

\todo{Type a table for acknowledgment so that I would have to write
``I thank'' so many times here? }Many people in both departments
of Mathematics and Computer Science need to do something more to make
my double-major works smoothly. I need to thank all of them, especially
directors of graduate study in Mathematics, especially Prof.\ Christopher
Judge and Prof.\ Matthias Weber; directors of Ph.D. study in Computer
Science, especially Prof.\ Yuqing Melanie Wu (吴愈青) and Prof.\ \textturkish{Funda
Ergun}.

@@

Thank all my friends for your support, including:

@@

\item Hao-Chun Lee (李浩君) and Hsien-Ching Kao (高憲慶) for discussing
what I did in the perspective of a general PhD. in Physics and working
later in more computational industry.

@@

\begin{abstract*}Most quantum computing models are based on the continuum
of real numbers, while classical digital computers faithfully realize
only discrete computational models. Analog computers appear to be
an option, but in reality are far weaker than would be needed for
computational models requiring real numbers. One approach to resolving
this conflict is to find consistent mathematical ways to limit measurement
precision to computable contexts that do not require incomputable
real numbers. Our goal is to build more philosophically consistent
models by investigating discrete quantum computing using finite number
systems, and, alternatively, by incorporating finite precision measurement
using intervals into quantum theory.

@@

Finally, we shift our attention to consider quantum interval-valued
probability measures (IVPMs), which potentially embody both finite
precision measurement and a sensible correspondence to standard quantum
probability. This interval-valued framework not only provides a natural
generalization of both classical IVPMs and conventional quantum probability
measures but also allows us to establish bounds on the validity of
the Kochen-Specker and Gleason theorems in realistic experimental
environments.\end{abstract*}

@@

Since a human cannot distinguish quantities differing only by an arbitrarily
small extent \cite{Turing_1937,Gisin2017}, we want to incorporate
this limitation into our quantum mechanical model to better capture
our ability to predict quantum phenomenon and the power of realistic
quantum computers. Since we are agnostic about whether the reality
is ultimately discrete or continuous, we considered two types of quantum
models. By assuming the quantum states are ultimately discrete and
distinguishable, our first type of quantum theories replaces the complex
numbers by finite fields. We then focus on the limitation of distinguishability
itself, incorporate the finite precision measurement into our model,
and consider quantum interval-valued probability measures.

After first suggested by Richard Feynman 40 years ago~\cite{Feynman1982Simulating},
IBM, \textitalian{Rigetti}, Google, Microsoft, and many other companies
and countries start to build realistic quantum computers recently
\cite{Shieber2018,Kahn2019} and seek more efficient ways to simulate
chemical molecules and to break down RSA \cite{boneh1999twenty,wiki:RSAFactoring}.
No matter how to, classical or quantum, attack these two tasks, they
require two different types of quantities. On one hand, a question
and its answer of RSA are both discrete integers. On the other hand,
molecules are described by quantities including the domain of their
position wave functions and the probability amplitudes of these wave
functions, which look like continuous at first sight. However, either
these quantities might later be discovered to be discrete, or their
exact values may never be pinned down. Hence, modeling quantum computing
based on the exact values of continuous quantities cannot really describe
the computational power of a realistic quantum computer. To better
understand this issue without struggling with the quantum theory,
it is instructive to review how classical quantities are characterized
and computed.

@@

In contrast, the non-apparently discrete quantities confused many
cultures since the very beginning. For example, the ancient Chinese
philosopher Chuang-tzu (庄子) didn't believe these quantities have a
minimum unit and expressed ``If a one-foot-long stick is cut into
halves every day, the cutting will never come to an end''\footnote{``一尺之棰，日取其半，萬世不竭。'' in Chinese \cite{Zhuangzi1999,Liu2018}.}.
As another example, Greek mathematicians in the school of Pythagoras
believed every mathematical model must be ultimately discrete, and
they in legend murdered Hippasus because he proved there is no minimum
unit between a side of a square and its diagonal. Comparing the idea
that the length of a square's diagonal is not a physical quantity
or there is no square physically, working with a mathematical model
without minimum unit seems more appealing. Since the mathematical
models are continuous, their corresponding physical quantities are
naturally assumed to be continuous. Later, these continuous quantities
become the basic component of analog computers. After logarithm was
discovered, a slide rule, also known as a slipstick, is used to compute
multiplication, division, and more complex operations as a mechanical
analog computer~\cite{wiki:SlideRule}. Charge amplifiers are used
to build electrical analog computers and compute the integration in
calculus \cite{wiki:Integrator,wiki:ChargeAmplifier}.

Although it is convenient to assume physical quantities is as continuous
as their mathematical models, each physical quantity discussed in
the previous paragraph could not have a simple one-to-one correspondence
with a real number. Chuang-tzu's stick is made by molecules and cannot
be cut in halves endlessly. The input charge of an integrator must
be an integer multiple of the elementary charge. A perfect square
cannot be drawn physically. Even we don't draw it, but merely ask
whether it exists in the space. This question is still beyond our
ability to answer because whether space itself is ultimately discrete
or continuous is still an unsolved question. Even if a perfect square
exists physically, we may not able to decide whether the ``square''
we discussed is really perfect or almost perfect differing only by
an arbitrarily small extent. Similar situation applying to the length
of a slide rule, it is beyond our ability to decide whether its length
is ultimately discrete or continuous. Even if its length is continuous,
identifying which real number represents its length requires infinite
precision, but the precision, in reality, is limited generally three
or four significant figures~\cite{wiki:AnalogComputer}. As the results,
even if analog computers really store real numbers, they cannot be
precisely read, written, and used to branch the computation. The opposite
of the last statement is sometimes assumed by real computational models.
For example, Blum, Cucker, Shub, and Smale modeled their BCSS machine
to allow the operations of deciding whether a number is greater or
equal to zero over real numbers and whether a number is exactly zero
over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Hence, when the quantities used to branch the computation closes to
zero, the branch chosen by a BCSS machine cannot reliably predict
the branch chosen by a realistic analog computer. Since two branches
of computations may not have any relations, the follow-up computational
paths may be significantly different, and this difference is unlikely
to be compensated by error analysis techniques easily like the butterfly
effect. Therefore, the computational power of analog computers might
be far weaker than the prediction based on this kind of theoretical
models.%
\begin{comment}
\todo{Maybe extended to consider probabilistic model? }
\end{comment}

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How does the theorical model predict the behavior of the Physical device? & 
Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Like its classical counterparts, quantum circuit model, the
most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states. To address the similar inconsistency, we first review the
``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}, where the probability amplitudes are in the field of complex numbers.
If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, considering the precision
of quantities could still result in a better model. However, comparing
to the precision of probability amplitudes, it is easier to consider
the precision of their inducing probabilities because the idea of
``imprecise probabilities'' has well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }This idea will be extended
to the quantum domain as quantum interval-valued probability measures
(QIVPM).

@@

Given a pure state~$\ket{\Phi}$, when we measure an observable represented
by a Hermitian matrix~$\mathbf{O}$, the measurement result is one
of the eigenvalues of $\mathbf{O}$. The probability of getting a
particular eigenvalue~$\lambda$ is $\melem{\Phi}{P}{\Phi}$, where
$P$ is the projection operator onto the eigenspace of $\lambda$.
This rule of computing the probability is called the Born rule \cite{Born1983,Mermin2007,Jaeger2007},
which is used when we want to extract information from a quantum computer.
For any mixed state~$\rho$, the generalized Born rule induces a
conventional quantum probability measure $\muB_{\rho}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projection operators on a given
Hilbert space. Conversely, any quantum probability measure $\mu:\events\rightarrow[0,1]$
can be induced from a mixed state $\rho$ in the Hilbert space of
dimension $D\ge3$ according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
In other word, this state $\rho$ is the unique state consistent with
any given quantum probability measure.
\end{comment}
The geometrical structure of states in CQT is first reviewed in Sec.~\ref{CQC1qubitBloch.sec},
where a pure state is represented as a vector in a Hilbert space,
and an irreducible state is actually a point in a projective Hilbert
space \cite{MosseriDandoloff2001,Jaeger2007} also known as a complex
projective space \cite{Hatcher2001,Bengtsson2007}. In Sec.~\ref{sec:fuzzy},
we will construct a quantum probability space based on classical probability
spaces, apply Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}
to recover the Born rule \cite{Born1983,Mermin2007,Jaeger2007}, and
define the expectation value of an observable. The expectation value
will then be used in Sec.~\ref{sec:The-geometry-of} to define purity
which provides an easy entanglement test.

\begin{comment}
In Section~\ref{modalquantum}, we examine previously-introduced
quantum theories defined over unrestricted finite fields and show
in Section~\ref{modalquantumcomputing} that this approach leads
to theories with such bizarre powers that they are probably nonphysical.
Although a version of quantum theory defined over a two-valued field
can express simple algorithms such as quantum teleportation, it is
so weak that it cannot express Deutsch's algorithm. This quantum theory
is, however, also so powerful that it can be used to solve an unstructured
database search of size $N$ using $O(\log(N))$ steps, which outperforms
the known asymptotic bound $O(\sqrt{N})$ in conventional quantum
computing.
\end{comment}
\begin{comment}
Our first discrete model replaces the complex numbers with an unrestricted
finite field~$\mathbb{F}_{q}$ \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
In this model, an $n$-qubit state is a non-zero vector in $\mathbb{F}_{q}^{2^{n}}$.
When we measure a given state~$\ket{\Phi}$, an observable is replaced
by a basis~$\mathcal{B}$ of $\mathbb{F}_{q}^{2^{n}}$. Since $\ket{\Phi}$
can be represented by the summation across a unique subset~$\mathcal{S}$
of the basis~$\mathcal{B}$, whether it is possible or impossible
to measure a basis vector~$\ket{i}$ depends on whether $\ket{i}$
is in $\mathcal{S}$ or not. Because this model only predicts whether
a result is possible or impossible, it is called the modal quantum
theory. Its computational model, called the modal quantum computing,
is far from conventional quantum computing. Although modal quantum
computing can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007}.
This quantum theory is, however, also so powerful that it can be used
to solve an unstructured database search of size $N$ using $O(\log(N))$
steps, which outperforms the known asymptotic bound $O(\sqrt{N})$
in conventional quantum computing \cite{Grover:1996:FQM:237814.237866,BennettBernsteinBrassardVazirani1997,Mermin2007,Jaeger2007}.
\end{comment}
In Chapter~\ref{chap:Unrestricted-Finite-Fields}, we examine previously-introduced
quantum theories defined over unrestricted finite fields \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
It is called the modal quantum theory because it can only predict
whether a measurement result is possible or impossible, but not its
probability. Although a version of a quantum theory defined over the
two-valued field can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007,Schumacher2012-SCHMQT},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007,DQT2014}.
This quantum theory is, however, also so powerful that it can be used
to efficiently solve the $\usat$ problem~\cite{Papadimitriou1993},
which is as hard as a general $\NP$-complete problem~\cite{Valiant198685}.
